Beyond Ultra-Diffuse Galaxies I: Mass-Size Outliers Among the Satellites of Milky Way Analogs

### 2022-10-25 paper1.tex
Updates from Overleaf

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
Large diffuse galaxies are hard to find, but understanding the environments where they live, their numbers, and ultimately their origins, is of intense interest and importance for galaxy formation and evolution. Using Subaru's Hyper Suprime-Cam Strategic Survey Program, we perform a systematic search for low surface brightness galaxies and present novel and effective methods for detecting and modeling them. 
As a case study, we build a large sample of galaxies that are outliers in the mass-size relation and are satellites of Milky Way analogs in the nearby Universe ($0.01 < z < 0.04$). These ``ultra-puffy'' galaxies (UPGs), defined to be $1.5\sigma$ above the average mass-size relation, represent the tail of the satellite size distribution. We find that each MW analog hosts $N_{\rm UPG} = 0.61\pm 0.04$ ultra-puffy galaxies on average, which is consistent with the observed abundance at this halo mass in the Local Volume. We also construct a sample of ultra-diffuse galaxies (UDGs) in MW analogs and find an abundance of $N_{\rm UDG} = 0.76\pm0.04$ per host. With literature results, we confirm that the UDG abundance scales with the host halo mass following a sublinear power law. We argue that our definition for ultra-puffy galaxies, which is based on the mass-size relation, is more physically-motivated than the common definition of ultra-diffuse galaxies, which depends on surface brightness and size cuts and thus yields different surface mass density cuts for quenched and star-forming galaxies. 
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
In a variety of galaxies in the Universe, there is an interesting subset of galaxies with low surface brightness \citep[dubbed low surface brightness galaxies or LSBGs, e.g.,][]{Sandage1984,Caldwell1987,Impey1988,McGaugh1995,Dalcanton1997a}. Recently, new excitement has been generated by the discovery of a large population of ultra-diffuse galaxies (UDGs) in the Coma cluster \citep{vanDokkum2015}, accompanied by many other works on detecting UDGs in clusters \citep[e.g.,][]{Koda2015,Mihos2015,Yagi2016,vdBurg2016,vdBurg2017,Lee2017,ManceraPina2018,Zaritsky2019,Danieli2019}, groups \citep[e.g.,][]{Roman2017b,Greco2018,SAGA-II,CarlstenELVES2022}, and the field \citep[e.g.,][]{Leisman2017,Roman2019,Prole2019,Tanoglidis2021,Kadowaki2021}. There have been a variety of studies of these diffuse galaxy systems, ranging from their dark matter content \citep[e.g.,][]{Mowla2017,vanDokkum2018,vanDokkum2019,Danieli2019DF2,Wasserman2019,Keim2022}, to their globular cluster populations \citep[e.g.,][]{vanDokkum2017,Somalwar2020,Forbes2020,Danieli2022,Gannon2022,vanDokkum2022GC}, and their stellar populations \citep[e.g.,][]{Gu2018,Ferre-Mateu2018,Pandya2018,Villaume2022}. We refer to LSBGs as objects selected purely by surface brightness without distance information. When knowing distances, people define UDGs to be galaxies with large sizes ($r_e > 1.5$ kpc) and low surface brightnesses ($\sbcen > 24.0\ \sbunit$), and thus UDGs are outliers in size and surface brightness compared to normal dwarf galaxies. A central question about LSBGs, including UDGs, is what causes them to be outliers from the average mass-size relation of dwarf galaxies.

@@

After applying the stellar mass and redshift cuts to the NSA catalog (\S\ref{sec:data}), there are 23,218 MW-like galaxies. We match them to the initial LSBG sample (described in \S \ref{sec:detection}, before the deblending step) as follows. For a given MW-like galaxy, we first calculate a virial mass assuming the stellar-to-halo mass relation in \citet{Behroozi2010} then calculate the corresponding virial radius $R_{\rm vir}$. It turns out that 40\% of our hosts have virial radii larger than 300 kpc, which is commonly used for the virial radius of the Milky Way. Then we identify any LSBG that falls inside the projected virial radius of a host. If one LSBG is matched to multiple hosts, we assign it to the nearest host based on the separation normalized by host virial radius. There are 901 MW-like hosts and 10,579 LSBG candidates associated with them.

% However, a significant fraction of spurious objects remain in this cross-matched LSBG sample, including galactic cirrus, tidal streams, shredded large galaxies, and compact sources blended with galaxy outskirts. 
As described in \S\ref{sec:deblending}, we then perform a deblending step to effectively remove spurious objects. There are 2,673 objects left after the deblending cuts. Next, we model these remaining objects using the Spergel profile as described in \S\ref{sec:modeling} and obtain their photometry and structural parameters from the best-fit models. We also remove duplicate objects. In the final catalog, we are left with 2,510 LSBG candidates associated with 689 MW analogs as our final LSBG sample. These MW analogs occupy a sky area of 75.94 deg$^{2}$ (out to $1\ R_{\rm vir}$). The host number drops to 689 because many hosts in the prior sample has no true positives as their satellites. We note that these LSBGs are matched to MW hosts in projection but do not have direct distance measurements, and we have not applied any statistical background correction at this stage.

@@

We randomly select a continuous patch of sky of 24 deg$^{2}$ in HSC PDR2 regardless of whether it contains MW analogs\footnote{$345\ \deg < \rm{R.A.} < 351\ \deg$, $-1.5\ \deg < \rm{Dec} < 2.5\ \deg$, which is not dominated by Galactic cirrus and stars. Such a patch of sky with 24 deg$^2$ is already enough for us to derive a background contamination fraction with small Poisson noise.}. Then we perform the same deblending and modeling steps for the 2,707 LSBG candidates detected in this region. We also remove objects that are already in our mass-size outlier samples and objects with completeness less than 0.1. We also did the same visual inspection as for UDGs and UPGs to remove false positives. In the end, there are 480 LSBGs (excluding false positives) representing a population of possible contaminants for the UDG and UPG samples. Next, because both UDG and UPG are defined based on a physical size, we randomly associate the 480 LSBGs with the 689 MW analogs that are surveyed in this paper. Then we calculate a physical size for each LSBG and classify an LSBG as a ``fake`` UDG or UPG if it satisfies the corresponding definition. We repeat such random matching 200 times. In the end, we obtain 8,392 fake UDGs and 8,485 fake UPGs. These UDGs and UPGs are ``fake'' only in the sense of being associated with random MW hosts.

In this way, the number density of fake UDGs is estimated to be $S_{\rm UDG} = 1.75\pm0.03\ \mathrm{deg}^{-2}$, and the number density of fake UPGs is $S_{\rm UPG} = 1.70\pm0.03\ \mathrm{deg}^{-2}$. The catalogs of fake UDG and UPGs are also available online\footnote{\url{https://astrojacobli.github.io/research/BeyondUDG/}}. Using the number densities and the probed area of our survey, we calculate the contamination fraction for both the UDG and UPG samples. For the UDG sample, we find $f_{\rm contam}^{\rm UDG} \approx 32\%$; for the UPG sample, the contamination fraction is $f_{\rm contam}^{\rm UPG} \approx 38\%$. We take the contamination fraction into account when calculating the abundances in \S\ref{sec:results}.

% For the UDG sample, we find $f_{\rm contam}^{\rm UDG} \approx 36\%$; for the UPG sample, the contamination fraction is $f_{\rm contam}^{\rm UPG} \approx 47\%$. We take the contamination fraction into account when calculating the abundances in \S\ref{sec:results}.

@@

\begin{figure*}
	\vbox{ 
		\centering
		\includegraphics[width=1\linewidth]{N_UDG_host_mass.pdf}
	}
    \caption{The abundance of UDGs as a function of host halo mass. \textit{Left}: We compile the UDG abundance measurements from the literature covering a wide range of host mass. The UDG abundance from our work $N_{\rm UDG} = 0.76\pm 0.04$ is shown as a red square, the power-law relation from \citet{vdBurg2017} is shown in gray, and our fitting result is shown in pink. Our UDG abundance is consistent with ELVES but significantly higher than that in SAGA. After including more measurements at the lower halo mass end, the power-law is shallower than the relation reported in \citet{vdBurg2017}. \textit{Right}: We split the UDG sample into bins based on the host halo mass (circles) and $g-i$ color (diamonds). The UDG abundance is higher for more massive host and redder host. }
    \label{fig:n_udg}
\end{figure*}

@@

We define the UDG (UPG) abundance as the average number of UDGs (UPGs) per host galaxy. We searched 689 MW analogs. In our UDG sample, there are 412 UDGs associated with 258 hosts. After correcting for background contamination (see \S \ref{sec:bkg}) and completeness (see \S \ref{sec:comp_meas}), the UDG abundance of MW analogs is $N_{\rm UDG} = 0.61\pm 0.04$ per host. The UPG abundance in our sample is $N_{\rm UPG} = 0.47\pm 0.04$ per host. Here we neglect the fact that the number of satellites contained within the virial sphere is different from the number of satellite within a projected cylinder with virial radius (the so-called deprojection factor, \citealt{vdBurg2017}).

We compare our UDG abundance with other surveys focused on MW analogs. In the SAGA survey \citep{SAGA-II}, 6 satellite galaxies (out of 127 spectroscopy-confirmed satellites around 36 MW analogs) satisfy the definition of UDG. Therefore the UDG abundance in \citet{SAGA-II} is about $N_{\rm UDG}^{\rm SAGA}\approx 0.17\pm0.07$. In the ELVES survey \citep{CarlstenELVES2022}, there are 13 UDGs out of 351 satellites with secure distances ($P_{\rm sat} > 0.8$) in 30 MW analogs, leading to a UDG abundance of $N_{\rm UDG}^{\rm ELVES} \approx 0.43\pm0.12$. If we only take UDGs more massive than our detection limit $M_\star > 10^{6.5}\ M_\odot$, we obtain slightly lower UDG abundances of $N_{\rm UDG, 6.5}^{\rm ELVES} \approx 0.36\pm0.11$. \citet{Roman2017b} identified 11 UDGs around three galaxy groups in the IAC Stripe 82 Legacy Survey \citep{Fliri2016}. Because both SAGA and ELVES select satellites based on distance measurements and do not correct for completeness, we interpret their UDG abundances as lower limits. We note that \citet{Roman2017b} do not estimate the background contamination fraction and do correct for completeness.

Because there is little literature on UPGs, we only compare our UPG abundance with ELVES. In ELVES, we take the 351 satellites with secure distances in 30 MW analogs and identify 36 UPGs associated with 20 hosts. Thus the UPG abundance fraction in ELVES is $N_{\rm UPG}^{\rm ELVES} = 1.20 \pm 0.20$. When only selecting UPGs with $M_\star > 10^{6.5}\ M_\odot$, we obtain 23 UPGs in 30 MW analogs, thus $N_{\rm UPG, 6.5}^{\rm ELVES} = 0.77 \pm 0.16$, which is slightly higher than our UPG abundance, but is consistent our result within $1\sigma$.

We plot the UDG abundances of these surveys in the left panel of Figure \ref{fig:n_udg}, together with the results for larger galaxy groups and clusters \citep{Koda2015,Munoz2015,Roman2017a,Roman2017b,Janssens2017,vdBurg2017}. 
The power-law regression result from \citet{vdBurg2017} $N_{\rm UDG} \propto M_h^{1.11\pm 0.07}$ is shown in gray. The UDG abundance of this work is highlighted as the red square. As shown in Figure \ref{fig:n_udg}, the scatter in the $N_{\rm UDG}-M_h$ relation gets larger at the lower halo-mass end, but this might be due to the small statistics in prior studies and the differences in completeness. Our UDG abundance is similar to the value in ELVES survey, but much higher than the value from SAGA. This might be explained by the different photometric depths of data used. As pointed out by \citet{CarlstenELVES2022,Font2022}, the SAGA survey reaches a surface brightness depth of $\sbeffr\approx 25\ \sbunit$, which is $\sim 2.5\ \sbunit$ shallower than the ELVES survey and this work. It is probable that SAGA missed some fraction of red and lower surface brightness galaxies. Our UDG abundance is marginally higher than the prediction from \citet{vdBurg2017} but is still consistent considering the large scatter of the power law.

Taking all data points from the literature (as shown in Figure \ref{fig:n_udg}), we use the orthogonal distance regression (ODR) to fit a power-law between $N_{\rm UDG}$ and host halo mass $M_h$ and find $N_{\rm UDG} \propto M_h^{0.96\pm 0.04}$, shown as the red solid line in Figure \ref{fig:n_udg}. The slope of the power law is shallower than \citet{vdBurg2017} ($\beta=1.11\pm0.07$) but steeper than \citet{Roman2017b} ($\beta=0.85\pm0.05$). \citet{ManceraPina2018} survey eight clusters and find a sublinear power law. Recently \citet{Karunakaran2022b} performed a similar analysis including SAGA, ELVES, and \citet{Nashimoto2022} and find a slightly shallower power law of $\beta=0.87\pm0.07$. We note that their surface brightness cut for UDGs is brighter than ours and thus they include more ``UDGs'' for MW-like hosts and have a shallower power law. One should be aware of the fact that literature results are not completeness corrected and background subtracted, and UDG definitions also vary across different studies. These effects could significantly bias the results when combining them. Therefore we caution that the power law index here should not be over-interpreted.

In the right panel of Figure \ref{fig:n_udg}, we split the UDG sample into different bins based on their host halo mass (shown as circles) and host color (shown as diamonds). We find that the UDG abundance is slightly higher for hosts with higher halo mass and redder color, but the trend is stronger with host color. 
\vspace{1em}

Another interesting quantity is the fraction of satellites that are mass-size outliers in a group or cluster. This fraction represents how efficient an environment is at producing puffy satellites. We denote this quantity as the UDG (UPG) fraction hereafter, which characterizes the tail of the size distribution of satellites. In order to calculate this fraction, we assume that MW analogs have $\sim 6 \pm 1$ satellites with $M_\star > 10^{6.5}\ M_\odot$ \citep{CarlstenELVES2022}. This lower limit in stellar mass is roughly our detection limit (see Fig. \ref{fig:mass_size}). Then for our samples, the UDG (UPG) fraction is $f_{\rm UDG} \approx 0.10 \pm 0.03$ and $f_{\rm UPG} \approx 0.08 \pm 0.03$. In ELVES, among 351 satellites with secure distances in 30 MW analogs, we identify 13 UDGs associated with 10 hosts, and 36 UPGs associated with 20 hosts. Thus the UDG fraction in ELVES is $f_{\rm UDG}^{\rm ELVES} \approx 0.04 \pm 0.01$, the UPG fraction in ELVES is $f_{\rm UPG}^{\rm ELVES} \approx 0.10 \pm 0.02$. We further select ELVES galaxies with $M_\star > 10^{6.5}\ M_\odot$ and find the same UDG and UPG fractions. We also compare our results with results in the Virgo cluster. Taking the data from the Next Generation Virgo Cluster Survey (NGVS, \citealt{Ferrarese2020}) and only considering satellites with $M_\star > 10^{6.5}\ M_\odot$, there are 14 UDGs and 17 UPGs out of 218 identified satellites (with no completeness correction). Thus their UDG (UPG) fraction is $f_{\rm UDG}^{\rm Virgo} \approx 0.06\pm 0.02$ and $f_{\rm UPG}^{\rm Virgo} \approx 0.08\pm 0.02$. However, we note that \citet{Ferrarese2020} only survey the central 4 deg$^2$ of the Virgo cluster, which probably biases the UDG (UPG) fraction. It seems that our UDG abundance and fraction are higher than both ELVES and NGVS. Our UPG abundance and fraction are consistent with ELVES and NVGS. In the future, a systematic and careful comparison on UDG (UPG) fraction in different environments is needed.

@@

However, in order to construct a UPG sample, one needs to know the distance, the color, a color-$M_\star/L$ relation to convert observed magnitudes to stellar mass, and a mass-size relation and the scatter therein. None of these are easy to obtain or free from systematic errors. Both size and stellar mass could have large uncertainties in the LSB regime. One also needs to apply a surface brightness cut (such as $\sbeff > 23.0\ \sbunit$ in this work) to make sure that the UPG sample is not contaminated by disk galaxies with high surface brightness due to the bulges.

@@

    \item After correcting for background contamination and completeness, the UDG abundance in MW analogs is $N_{\rm UDG} = 0.76\pm 0.04$ per host and the UPG abundance is $N_{\rm UPG} = 0.61 \pm 0.04$ per host. Our UDG abundance is slightly higher than that in ELVES, but much higher than that in SAGA. When considering the detection depth, our UPG abundance agrees with ELVES result within $1\sigma$.

@@

    \item For a MW analog, on average, about 12\% of its satellites are UDGs and about 10\% of its satellites are UPGs. Although the UPG fraction is consistent with ELVES and NGVS, the UDG fraction is found to be higher than those in ELVES and NGVS.

@@

\section*{Acknowledgment}
JL is grateful for discussions with Meng Gu, ChangHoon Hahn, Sihao Cheng. The authors thank Yao-Yuan Mao for his \href{https://github.com/yymao/decals-image-list-tool}{tool} for visualizing the image cutouts. JEG gratefully acknowledges support from NSF grant AST-1007052. J.H.K acknowledges the support from the National Research Foundation of Korea (NRF) grant, No. 2021M3F7A1084525, funded by the Korea government (MSIT).

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
Large diffuse galaxies are hard to find, but understanding the environments where they live, their numbers, and ultimately their origins, is of intense interest and importance for galaxy formation and evolution. Using Subaru's Hyper Suprime-Cam Strategic Survey Program, we perform a systematic search for low surface brightness galaxies and present novel and effective methods for detecting and modeling them. 
As a case study, we build a large sample of galaxies that are outliers in the mass-size relation and are satellites of Milky Way analogs in the nearby Universe ($0.01 < z < 0.04$). These ``ultra-puffy'' galaxies (UPGs), defined to be $1.5\sigma$ above the average mass-size relation, represent the tail of the satellite size distribution. We find that each MW analog hosts $N_{\rm UPG} = 0.31\pm 0.05$ ultra-puffy galaxies on average, which is consistent but slightly lower than the observed abundance at this halo mass in the Local Volume. We also construct a sample of ultra-diffuse galaxies (UDGs) in MW analogs and find an abundance of $N_{\rm UDG} = 0.44\pm0.05$ per host. With literature results, we confirm that the UDG abundance scales with the host halo mass following a sublinear power law. We argue that our definition for ultra-puffy galaxies, which is based on the mass-size relation, is more physically-motivated than the common definition of ultra-diffuse galaxies, which depends on surface brightness and size cuts and thus yields different surface mass density cuts for quenched and star-forming galaxies. 
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
In a variety of galaxies in the Universe, there is an interesting subset of galaxies with low surface brightness \citep[dubbed low surface brightness galaxies or LSBGs, e.g.,][]{Sandage1984,Caldwell1987,Impey1988,McGaugh1995,Dalcanton1997a}. Recently, new excitement has been generated by the discovery of a large population of ultra-diffuse galaxies (UDGs) in the Coma cluster \citep{vanDokkum2015}, accompanied by many other works on detecting UDGs in clusters \citep[e.g.,][]{Koda2015,Mihos2015,Yagi2016,vdBurg2016,vdBurg2017,Lee2017,ManceraPina2018,Zaritsky2019,Danieli2019}, groups \citep[e.g.,][]{Roman2017b,Greco2018,SAGA-II,Roman2021,CarlstenELVES2022}, and the field \citep[e.g.,][]{Leisman2017,Roman2019,Prole2019,Tanoglidis2021,Kadowaki2021}. There have been a variety of studies of these diffuse galaxy systems, ranging from their dark matter content \citep[e.g.,][]{Mowla2017,vanDokkum2018,vanDokkum2019,Danieli2019DF2,Wasserman2019,Keim2022}, to their globular cluster populations \citep[e.g.,][]{vanDokkum2017,Somalwar2020,Forbes2020,Danieli2022,Gannon2022,vanDokkum2022GC}, and their stellar populations \citep[e.g.,][]{Gu2018,Ferre-Mateu2018,Pandya2018,Villaume2022}. We refer to LSBGs as objects selected purely by surface brightness without distance information. When knowing distances, people define UDGs to be galaxies with large sizes ($r_e > 1.5$ kpc) and low surface brightnesses ($\sbcen > 24.0\ \sbunit$), and thus UDGs are outliers in size and surface brightness compared to normal dwarf galaxies. A central question about LSBGs, including UDGs, is what causes them to be outliers from the average mass-size relation of dwarf galaxies.

@@

After applying the stellar mass and redshift cuts to the NSA catalog (\S\ref{sec:data}), there are 23,218 MW-like galaxies. We then select those within HSC PDR2 footprint and match them to the initial LSBG sample (described in \S \ref{sec:detection}, before the deblending step) as follows. For a given MW-like galaxy, we first calculate a virial mass assuming the stellar-to-halo mass relation in \citet{Behroozi2010} then calculate the corresponding virial radius $R_{\rm vir}$. It turns out that 40\% of our hosts have virial radii larger than 300 kpc, which is commonly used for the virial radius of the Milky Way. Then we identify any LSBG that falls inside the projected virial radius of a host. If one LSBG is matched to multiple hosts, we assign it to the nearest host based on the separation normalized by host virial radius. There are 922 MW-like hosts and 10,579 LSBG candidates associated with them. These MW analogs occupy a sky area of 89.19 deg$^{2}$ (out to $1\ R_{\rm vir}$).

% However, a significant fraction of spurious objects remain in this cross-matched LSBG sample, including galactic cirrus, tidal streams, shredded large galaxies, and compact sources blended with galaxy outskirts. 
As described in \S\ref{sec:deblending}, we perform a deblending step to effectively remove spurious objects. There are 2,673 objects left after the deblending cuts. Next, we model these remaining objects using the Spergel profile as described in \S\ref{sec:modeling} and obtain their photometry and structural parameters from the best-fit models. We also remove duplicate objects. In the final catalog, we are left with 2,510 LSBG candidates associated with 689 MW analogs as our final LSBG sample. The host number drops from 922 to 689 because many hosts in the prior sample has no true positives as their satellites. We note that these LSBGs are matched to MW hosts in projection but do not have direct distance measurements, and we have not applied any statistical background correction at this stage.

@@

We randomly select a continuous patch of sky of 24 deg$^{2}$ in HSC PDR2 regardless of whether it contains MW analogs\footnote{$345\ \deg < \rm{R.A.} < 351\ \deg$, $-1.5\ \deg < \rm{Dec} < 2.5\ \deg$, which is not dominated by Galactic cirrus and stars. Such a patch of sky with 24 deg$^2$ is already enough for us to derive a background contamination fraction with small Poisson noise.}. Then we perform the same deblending and modeling steps for the 2,707 LSBG candidates detected in this region. We also remove objects that are already in our mass-size outlier samples and objects with completeness less than 0.1. We also did the same visual inspection as for UDGs and UPGs to remove false positives. In the end, there are 480 LSBGs (excluding false positives) representing a population of possible contaminants for the UDG and UPG samples. Next, because both UDG and UPG are defined based on a physical size, we randomly associate the 480 LSBGs with the 922 MW analogs that are surveyed in this paper. Then we calculate a physical size for each LSBG and classify an LSBG as a ``fake`` UDG or UPG if it satisfies the corresponding definition. We repeat such random matching 200 times. In the end, we obtain 7,625 fake UDGs and 8,267 fake UPGs. These UDGs and UPGs are ``fake'' only in the sense of being associated with random MW hosts.

In this way, the number density of fake UDGs is estimated to be $S_{\rm UDG} = 1.60\pm0.25\ \mathrm{deg}^{-2}$, and the number density of fake UPGs is $S_{\rm UPG} = 1.72\pm0.23\ \mathrm{deg}^{-2}$. The catalogs of fake UDG and UPGs are also available online\footnote{\url{https://astrojacobli.github.io/research/BeyondUDG/}}. Using the number densities and the probed area of our survey (89.19 deg$^{2}$ occupied by 922 hosts), we calculate the contamination fraction for both the UDG and UPG samples. For the UDG sample, we find $f_{\rm contam}^{\rm UDG} \approx 35\%\pm5\%$; for the UPG sample, the contamination fraction is $f_{\rm contam}^{\rm UPG} \approx 45\%\pm6\%$. We take the contamination fraction into account when calculating the abundances in \S\ref{sec:results}.

@@

\begin{figure*}
	\vbox{ 
		\centering
		\includegraphics[width=1\linewidth]{N_UDG_host_mass.pdf}
	}
    \caption{The abundance of UDGs as a function of host halo mass. \textit{Left}: We compile the UDG abundance measurements from the literature covering a wide range of host mass. The UDG abundance from our work $N_{\rm UDG} = 0.44\pm 0.05$ is shown as a red square, the power-law relation from \citet{vdBurg2017} is shown in gray, and our fitting result is shown in pink. Our UDG abundance is consistent with ELVES but significantly higher than that in SAGA. After including more measurements at the lower halo mass end, the power-law is shallower than the relation reported in \citet{vdBurg2017}. \textit{Right}: We split the UDG sample into bins based on the host halo mass (circles) and $g-i$ color (diamonds). The UDG abundance is higher for more massive host and redder host. }
    \label{fig:n_udg}
\end{figure*}

@@

We define the UDG (UPG) abundance as the average number of UDGs (UPGs) per host galaxy. We searched 922 MW analogs. In our UDG sample, there are 412 UDGs associated with 258 hosts. After correcting for background contamination (see \S \ref{sec:bkg}) and completeness (see \S \ref{sec:comp_meas}), the UDG abundance of MW analogs is $N_{\rm UDG} = 0.44\pm 0.05$ per host. The UPG abundance in our sample is $N_{\rm UPG} = 0.31\pm 0.04$ per host. Here we neglect the fact that the number of satellites contained within the virial sphere is different from the number of satellite within a projected cylinder with virial radius (the so-called deprojection factor, \citealt{vdBurg2017}).

We compare our UDG abundance with other surveys focused on MW analogs. In the SAGA survey \citep{SAGA-II}, 6 satellite galaxies (out of 127 spectroscopy-confirmed satellites around 36 MW analogs) satisfy the definition of UDG. Therefore the UDG abundance in \citet{SAGA-II} is about $N_{\rm UDG}^{\rm SAGA}\approx 0.17\pm0.07$. In the ELVES survey \citep{CarlstenELVES2022}, there are 13 UDGs out of 351 satellites with secure distances ($P_{\rm sat} > 0.8$) in 30 MW analogs, leading to a UDG abundance of $N_{\rm UDG}^{\rm ELVES} \approx 0.43\pm0.12$. If we only take UDGs more massive than our detection limit $M_\star > 10^{6.5}\ M_\odot$, we obtain slightly lower UDG abundances of $N_{\rm UDG, 6.5}^{\rm ELVES} \approx 0.36\pm0.11$. Because both SAGA and ELVES select satellites based on distance measurements and do not correct for completeness, we interpret their UDG abundances as lower limits. \citet{Roman2017b} identified 11 UDGs around three galaxy groups in the IAC Stripe 82 Legacy Survey \citep{Fliri2016}, but it is hard to calculate a UDG abundance due to small number statistics. Our UDG abundance is consistent with the value in ELVES survey, but higher than the value from SAGA. This might be explained by the different photometric depths of data used. As pointed out by \citet{CarlstenELVES2022,Font2022}, the SAGA survey reaches a surface brightness depth of $\sbeffr\approx 25\ \sbunit$, which is $\sim 2.5\ \sbunit$ shallower than the ELVES survey and this work. It is probable that SAGA missed some fraction of red and lower surface brightness galaxies. %We note that \citet{Roman2017b} do not estimate the background contamination fraction and do correct for completeness. \jiaxuan{Due to different methods, completeness, depth, comparisons can be hard}

We plot the UDG abundances from these surveys in the left panel of Figure \ref{fig:n_udg}, together with the results for larger galaxy groups and clusters \citep{Koda2015,Munoz2015,Roman2017a,Roman2017b,Janssens2017,vdBurg2017}. 
The power-law regression result from \citet{vdBurg2017} $N_{\rm UDG} \propto M_h^{1.11\pm 0.07}$ is shown in gray. The UDG abundance of this work is highlighted as the red square. As shown in Figure \ref{fig:n_udg}, the scatter in the $N_{\rm UDG}-M_h$ relation gets larger at the lower halo-mass end, but this might be due to the small statistics in prior studies and the differences in completeness. Our UDG abundance is marginally higher than the prediction from \citet{vdBurg2017} but is still consistent considering the large scatter of the power law.

Taking all data points from the literature (as shown in Figure \ref{fig:n_udg}), we use the orthogonal distance regression (ODR) to fit a power-law between $N_{\rm UDG}$ and host halo mass $M_h$ and find $N_{\rm UDG} \propto M_h^{0.96\pm 0.04}$, shown as the red solid line in Figure \ref{fig:n_udg}. The slope of the power law is shallower than \citet{vdBurg2017} ($\beta=1.11\pm0.07$) but steeper than \citet{Roman2017b} ($\beta=0.85\pm0.05$). \citet{ManceraPina2018} survey eight clusters and find a sublinear power law. Recently \citet{Karunakaran2022b} performed a similar analysis including SAGA, ELVES, and \citet{Nashimoto2022} and find a slightly shallower power law of $\beta=0.87\pm0.07$. We note that their surface brightness cut for UDGs is brighter than ours and thus they include more ``UDGs'' for MW-like hosts and have a shallower power law. One should be aware of the fact that literature results are of different depths, often not completeness corrected and background subtracted, and UDG definitions also vary across different studies. These effects could significantly bias the results when combining them. Therefore we caution that the UDG abundance and the power law index here should not be over-interpreted.

In the right panel of Figure \ref{fig:n_udg}, we split the UDG sample into different bins based on their host halo mass (shown as circles) and host color (shown as diamonds). We find that the UDG abundance is slightly higher for hosts with higher halo mass and redder color, but the trend is stronger with host color.

Because there is little literature on UPGs, we only compare our UPG abundance with ELVES. In ELVES, we take the 351 satellites with secure distances in 30 MW analogs and identify 36 UPGs associated with 20 hosts. Thus the UPG abundance fraction in ELVES is $N_{\rm UPG}^{\rm ELVES} = 1.20 \pm 0.20$. When only selecting UPGs with $M_\star > 10^{6.5}\ M_\odot$, we obtain 23 UPGs in 30 MW analogs, thus $N_{\rm UPG, 6.5}^{\rm ELVES} = 0.77 \pm 0.16$, which is slightly higher than our UPG abundance.

Another interesting quantity is the fraction of satellites that are mass-size outliers in a group or cluster. This fraction represents how efficient an environment is at producing puffy satellites. We denote this quantity as the UDG (UPG) fraction hereafter, which characterizes the tail of the size distribution of satellites. In order to calculate this fraction, we assume that MW analogs have $\sim 6 \pm 1$ satellites with $M_\star > 10^{6.5}\ M_\odot$ \citep{CarlstenELVES2022}. This lower limit in stellar mass is roughly our detection limit (see Fig. \ref{fig:mass_size}). Then for our samples, the UDG (UPG) fraction is $f_{\rm UDG} \approx 0.07 \pm 0.02$ and $f_{\rm UPG} \approx 0.05 \pm 0.02$. In ELVES, among 351 satellites with secure distances in 30 MW analogs, we identify 13 UDGs associated with 10 hosts, and 36 UPGs associated with 20 hosts. Thus the UDG fraction in ELVES is $f_{\rm UDG}^{\rm ELVES} \approx 0.04 \pm 0.01$, the UPG fraction in ELVES is $f_{\rm UPG}^{\rm ELVES} \approx 0.10 \pm 0.02$. We further only select ELVES galaxies with $M_\star > 10^{6.5}\ M_\odot$ and find the same UDG and UPG fractions. We also compare our results with results in the Virgo cluster. Taking the data from the Next Generation Virgo Cluster Survey (NGVS, \citealt{Ferrarese2020}) and only considering satellites with $M_\star > 10^{6.5}\ M_\odot$, there are 14 UDGs and 17 UPGs out of 218 identified satellites (with no completeness correction). Thus their UDG (UPG) fraction is $f_{\rm UDG}^{\rm Virgo} \approx 0.06\pm 0.02$ and $f_{\rm UPG}^{\rm Virgo} \approx 0.08\pm 0.02$. However, we note that \citet{Ferrarese2020} only survey the central 4 deg$^2$ of the Virgo cluster, which probably biases the UDG (UPG) fraction.
We interpret our UDG and UPG fractions to be consistent with ELVES and NGVS. However, we emphasize that it is challenging to compare across various searches due to the differences in the host sample selection, depth, completeness, and contamination estimation. In the future, a more systematic and careful comparison on UDG (UPG) fraction in different environments is needed.

@@

However, in order to construct a UPG sample, one needs to know the distance, the color, a color-$M_\star/L$ relation to convert observed magnitudes to stellar mass, and a mass-size relation and the scatter therein. None of these are easy to obtain or free from systematic errors. Both size and stellar mass could have large uncertainties in the LSB regime. One also needs to apply a surface brightness cut (such as $\sbeff > 23.0\ \sbunit$ in this work) to make sure that the UPG sample is not dominated by disk galaxies with high surface brightness due to the bulges. In \S\ref{sec:bkg}, we find that the UPG sample has a higher contamination fraction than the UDG sample. Since the UPG definition does not preferentially exclude blue LSBGs, background disk galaxies are more easily to be identified as UPGs. Based on our test, a UPG sample with more aggressive cut (e.g., $>2\sigma$ above the average mass-size relation) would have a higher purity, but only possible when one have a much larger LSBG sample.

@@

    \item After correcting for background contamination and completeness, the UDG abundance in MW analogs is $N_{\rm UDG} = 0.44\pm 0.05$ per host and the UPG abundance is $N_{\rm UPG} = 0.31 \pm 0.04$ per host. Our UDG abundance agrees with ELVES quite well, but is much higher than that in SAGA. We obtain a UPG abundance lower than that in ELVES.

@@

    \item For a MW analog, on average, about $7\%\pm2\%$ of its satellites are UDGs and about $5\%\pm2\%$ of its satellites are UPGs. Our UDG and UPG fraction are consistent with ELVES and NGVS.

@@

\section*{Acknowledgment}
JL is grateful for discussions with Meng Gu, ChangHoon Hahn, Sihao Cheng. The authors thank Yao-Yuan Mao for his \href{https://github.com/yymao/decals-image-list-tool}{tool} for visualizing the image cutouts. JEG gratefully acknowledges support from NSF grant AST-1007052. J.H.K acknowledges the support from the National Research Foundation of Korea (NRF) grant, No. 2021M3F7A1084525, funded by the Korea government (MSIT). S.D. is supported by NASA through Hubble Fellowship grant HST-HF2-51454.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
